---
title: "Class 11 hw"
author: "Lucy Wang"
format: html
---

# Section 4: Population Scale Analysis


> Q13: Read this file into R and determine the sample size for each genotype and their corresponding median expression levels for each of these genotypes.


Read in the data:
```{r}
expr <- read.table("rs8067378_ENSG00000172057.6.txt")
head(expr)
```

How many samples do we have - check the number of rows:
```{r}
nrow(expr)
```

```{r}
#Genotype info
table(expr$geno)
```

```{r}
aa <- expr$exp[expr$geno == "A/A"]
ag <- expr$exp[expr$geno == "A/G"]
gg <- expr$exp[expr$geno == "G/G"]
summary(aa)
summary(ag)
summary(gg)
```


```{r}
library(ggplot2)
```

make a boxplot w/ ggplot

```{r}
#use `notch` get a more clear indication
ggplot(expr) + aes(geno, exp, fill=geno) +
  geom_boxplot(notch = TRUE)
```

> Q14: Generate a boxplot with a box per genotype, what could you infer from the relative expression value between A/A and G/G displayed in this plot? Does the SNP effect the expression of ORMDL3?

A/A expression is higher than G/G expression, with almost no overlap between their central 50% (the box). It can be inferred that this SNP affect the expression of ORMDL3 from the data.

# THE END