---
title: "Class 5: Data Visualization"
author: "Lucy Wang"
format: html
---

# Plotting in R

R has multiple plotting and graphics systems. The most popular of which is **ggplot2**.

We have already played with "base" R graphics. This comes along with R "out of the box". (WoW)

```{r}
plot(cars)
```

Compared to base R plots ggplot is much more verbose - I need to write more code to get simple plots like the above.

To use ggplot I need to first install the ggplot2 package. To install any package in R, I use the `install.packges()` command along with the package name.

The install is a one time only requirement. The package is now on our computer. I don't need to re-install it.

However, I can't just use it without loading it up with a `library()` call.


```{r}
library(ggplot2)
```

```{r}
ggplot(cars)
```

All ggplot figures need at least 3 things:

- data (this is the `data.frame` with our numbers)
- aesthetics (`aes()`, how our data maps to the plot)
- geoms (do u want lines, points, columns, etc...)



```{r}
ggplot(data = cars) + 
  aes(x=speed, y=dist) +
  geom_point() +
  geom_line()

```

That is not we want, unfortunately.

```{r}
ggplot(data=cars) + 
  aes(x=speed, y=dist) +
  geom_point() +
  labs(title="Speed and Stopping Distances of Cars",
       x="Speed (MPH)", 
       y="Stopping Distance (ft)",
       subtitle = "Back in the Old Old Time",
       caption="Dataset: 'cars'") +
  geom_smooth(method = "lm", se=FALSE) +
  theme_bw()


```


```{r}
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"

genes <- read.delim(url)
head(genes)
```


The `head()` will only print the first 6 rows by default.

```{r}
nrow(genes)
```

```{r}
table(genes$State)
```

```{r}
round( table(genes$State)/nrow(genes) * 100, 2 )
```


```{r}
q <- ggplot(genes) +
  aes(Condition1, Condition2, color=State) +
  geom_point() +
  labs(title = "State Genes",
       subtitle = "Lots of them")

q
```


Can do **bold** and *italics*

```{r}
q + scale_colour_manual( values=c("blue","gray","red") ) +
    labs(title="Gene Expresion Changes Upon Drug Treatment",
         x="Control (no drug) ",
         y="Drug Treatment")
```


```{r}
url <- "https://raw.githubusercontent.com/jennybc/gapminder/master/inst/extdata/gapminder.tsv"

gapminder <- read.delim(url)
```

```{r}
library(dplyr)

gapminder_2007 <- gapminder %>% filter(year==2007)
```



```{r}
ggplot(gapminder_2007) +
  aes(x=gdpPercap, y=lifeExp, color=continent, size=pop)+
  geom_point(alpha=0.5)
```


```{r}
ggplot(gapminder_2007) + 
  aes(x = gdpPercap, y = lifeExp, color = pop) +
  geom_point(alpha=0.8)
```



```{r}
ggplot(gapminder_2007) + 
  geom_point(aes(x = gdpPercap, y = lifeExp,
                 size = pop), alpha=0.5) + 
  scale_size_area(max_size = 25)
```


```{r}
gapminder_1957 <- gapminder %>% filter(year==1957)

ggplot(gapminder_1957) + 
  aes(x = gdpPercap, y = lifeExp, color=continent,
                 size = pop) +
  geom_point(alpha=0.7) + 
  scale_size_area(max_size = 10) 
```



```{r}
gapminder_1957 <- gapminder %>% filter(year==1957 | year==2007)

ggplot(gapminder_1957) + 
  geom_point(aes(x = gdpPercap, y = lifeExp, color=continent,
                 size = pop), alpha=0.7) + 
  scale_size_area(max_size = 11) +
  facet_wrap(~year)
```

data-to-viz.com