---
title: "Class06 HW"
author: "Lucy Wang"
date: "2023-01-30"
output: html_document
---

---
title: "Class06 HW"
author: "Lucy Wang"
format: html
---

```{r}
# (A. Can you improve this analysis code?
df <- data.frame(a=1:10, b=seq(200,400,length=10),c=11:20,d=NA)
df$a <- (df$a - min(df$a)) / (max(df$a) - min(df$a))
df$b <- (df$b - min(df$a)) / (max(df$b) - min(df$b))
df$c <- (df$c - min(df$c)) / (max(df$c) - min(df$c))
df$d <- (df$d - min(df$d)) / (max(df$a) - min(df$d))
```

```{r}
#Improved code A)
#Naming all the input as "x"
D <- function(x){
  #Doing the exact same calculation thing as previous code does
  (x - min(x)) / (max(x) - min(x))
}
```

To apply the `D` function, simply put `D(data)` and press return. Or use the `apply()` function to call `D` . For example, one can do:

`apply(data, 1, D)`



```{r}
# Can you improve this analysis code?
library(bio3d)
s1 <- read.pdb("4AKE")  # kinase with drug
s2 <- read.pdb("1AKE")  # kinase no drug
s3 <- read.pdb("1E4Y")  # kinase with drug
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s1, chain="A", elety="CA")
s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```
**Q1. What type of object is returned from the read.pdb() function?**   
It returns a list of class "pdf"(from `?read.pdb`).

**Q2. What does the trim.pdb() function do?**   
It produces a new smaller PDB object, containing a subset of atoms, from a given larger PDB object (from `?trim.pdb`).

**Q3. What input parameter would turn off the marginal black and grey rectangles in the plots and what do they represent in this case?**   
I can set `top = FALSE` and `bot = FALSE` within `plotb3()` to turn the top and bottom rectangles off. They represent the optional secondary structure of the given data (from `?plotb3`).

```{r}
library(bio3d)
s1 <- read.pdb("4AKE")  # kinase with drug
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s1.b <- s1.chainA$atom$b
plotb3(s1.b, sse=s1.chainA, typ="l", top = FALSE, bot = FALSE, ylab="Bfactor")
```

**Q4. What would be a better plot to compare across the different proteins?**
Bar chart can be a better choice since it is discontinuous and allows to see seperate protein structures.

**Q5. Which proteins are more similar to each other in their B-factor trends. How could you quantify this?**.  
```{r}
hc <- hclust( dist( rbind(s1.b, s2.b, s3.b) ) )
plot(hc)
```
I use `rbind()` to combine a sequence of data by row, then use `dist()` to get a distance matrix computed from the data. In the end I use the     `hclust()` to compute the dissimilarity structure produced by `dist()`. After these I can finally plot the data.

```{r}
#Improved Code for Above
```



